NMRlipids III: Lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations

### 2019-05-07 MANUSCRIPT/manuscript.tex
Add separate supplementary information

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\begin{document}

@@

\subsection{Structural sampling of molecules in bilayers with cholesterol}
Order parameters for C-H bond vectors in lipid bilayer systems, measured
with $^{13}$C or $^{2}$H NMR techniques, give indirect information about the structural
sampling of indidual molecules \cite{ollila16}. The observed increase of
acyl chain order parameters with added cholesterol in simulations and
experiments can be explained by increased trans conformations in acyl chains \cite{ferreira13,??},
which is suggested to play critical role in the phase behaviour of PC-lipid--cholesterol
mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is
expected to be a necessary condition for a model used to understand lipid--cholesterol
phase behaviour.

Acyl chain order parameters for pure 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
and mixture with 50 mol\% of cholesterol from different simulations and experiments
are shown in Fig. \ref{OrderParametersCHOL}. Phase separation is not observed for this system \cite{ionova12,ferreira13},
thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
Experimental acyl chain order parameters are typically well reproduced in the state of the art lipid models
for pure lipid bilayers (for review see \cite{ollila16}). This is also mainly observed in the data
presented here, except that CHARMM36 slightly overestimates the order parameters and MACROG model
underestimates order parameters in the beginning of sn-1 chain. In equimolar mixtures with cholesterol
all models overestimate the order parameters. For Berger/H\"oltje and MACROG models the overestimation is significant,
while less severe for CHARMM36 and Slipid models.\\
\todo{Is overestimation in CHARMM36 and Slipid significant or not?} \\
\todo{Lipid14 results from recent article should be added in this figure} 
\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
  \caption{\label{OrderParametersCHOL}
    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
  }
\end{figure*}

The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
(currently only sn-1). \\
\todo{This figure should be finished and similar for sn-2 should be done as well}
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/OrderParametersCHANGEScholSN1.eps}
  \caption{\label{OrderParametersCHOLchanges}
    Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.}

@@

Simulated systems are listed in Table \ref{systemsCHOL} and the full simulation
details are given in references or in supplementary material.

@@

@@

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
% Put your acknowledgments here.
\end{acknowledgments}
\newpage
\appendix
\begin{center}
{\bf SUPPLEMENTARY INFORMATION}
\end{center}
\section{CHARMM36 results from different simulation packages}
The results from CHARMM36 model for lipid bilayers from different 
simulation packages have been reported to give different results in
the literature \cite{piggot12,lee16}. The results are mainly
dependent on different Lennard-Jones cut-off settings, but
all the details are not quite understood. In this work we use
the results from Gromacs 5 with settings suggested to be optimal
by Gromacs webpage.~\todo{List these settings here or in SI.} We also compared the results from Gromacs 5 with
these settings to the results simulated with NAMD, OpenMM and literature
values. Based on comparison shown in Fig. \ref{programsCOMP}, we conclude
that Gromacs 5 with settings suggested in webpage gives consistent
results with the literature and other simulation packages. Thus these
results are used in the main body of the paper. However, order parameters
are slightly overestimated respected to the experiments also with these
settings. \\
\todo{Should be extend this discussion based on discussion at  https://github.com/NMRLipids/NmrLipidsCholXray/issues/4 ?}
 \begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/OrderParametersPROGRAMS.eps}

  \caption{\label{programsCOMP}
    Results for CHARMM36 model \cite{klauda10} from different simulation packages.
    Discussion going on at {\it https://github.com/NMRLipids/NmrLipidsCholXray/issues/4}.
  }
\end{figure}

% Create the reference section using BibTe
\bibliography{refs.bib}

>>>>

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{xr}
\externaldocument{manuscriptSUPPL}
\begin{document}

@@

Order parameters for C-H bond vectors in lipid bilayer systems, measured
with $^{13}$C or $^{2}$H NMR techniques, give indirect information about the structural
sampling of indidual molecules \cite{ollila16}. The observed increase of
acyl chain order parameters with added cholesterol in simulations and
experiments can be explained by increased trans conformations in acyl chains \cite{ferreira13,??},
which is suggested to play critical role in the phase behaviour of PC-lipid--cholesterol
mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is
expected to be a necessary condition for a model used to understand lipid--cholesterol
phase behaviour.

@@

Molecular dynamics simulation data was collected using the open collaboration method \cite{botan15}.
Simulated systems are listed in Table \ref{systemsCHOL} and the full simulation
details are given in references or in supplementary material.

@@

\subsection{Monitoring molecular interactions between POPC and cholesterol in lipid bilayer using NMR and MD simulations}

Addition of cholesterol substantially increases the acyl chain C-H order parameters in PC lipid bilayers (Fig. \ref{OrderParametersCHOL}),
which can be explained by increased trans conformations in acyl chains \cite{ferreira13,??}.
%Acyl chain order parameters for pure 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
%and mixture with 50 mol\% of cholesterol from different simulations and experiments
%are shown in Fig. \ref{OrderParametersCHOL}.
The absolute values of order parameters of sn-1 chain from $^{13}$C NMR experiments \cite{ferreira13}
exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because
phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
%thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.

Experimental acyl chain order parameters are typically well reproduced in the state of the art lipid models
for pure lipid bilayers (for review see \cite{ollila16}). This is also mainly observed in the data
presented here, except that CHARMM36 slightly overestimates the order parameters and MACROG model
underestimates order parameters in the beginning of sn-1 chain. In equimolar mixtures with cholesterol
all models overestimate the order parameters. For Berger/H\"oltje and MACROG models the overestimation is significant,
while less severe for CHARMM36 and Slipid models.\\
\todo{Is overestimation in CHARMM36 and Slipid significant or not?} \\
\todo{Lipid14 results from recent article should be added in this figure} 
\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
  \caption{\label{OrderParametersCHOL}
    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
  }
\end{figure*}

The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
(currently only sn-1). \\
\todo{This figure should be finished and similar for sn-2 should be done as well}
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/slopes.eps}
  \caption{\label{OrderParametersCHOLchanges}
    Slopes of order parameters as a function of cholesterol.
    %Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.
  }

@@

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
% Put your acknowledgments here.
\end{acknowledgments}

\bibliography{refs.bib}
